# Set-up

Run these commands in the R terminal

install.packages("tidyverse")
install.packages("ggplot2")

Install these packages:

```{r}
library(tidyverse)
library(ggplot2)
```

This notebook is the final part of our project – visualisation. Using our final dataframe, we are able to visualise data through applications of R and ggplot2.

Important!!! Ensure the current working directory is in the main DS folder.

```{r}
getwd()
```

```{r}
df <- read.csv('../data/final_compiled.csv')
```

```{r}
head(df)
```

# Statistical inference

```{r}
selected_columns <- df[, c('like_count','view_count','comment_count', 'sentiment_positive', 'sentiment_neutral', 'sentiment_negative', 'sentiment_compound_absolute', 'lexical_richness', 'song_length', 'popularity')]

corr_df <- cor(selected_columns, use="complete.obs")
```

```{r}
corr_df2 <- corr_df %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "rowname") %>%  # Convert row names to a column
  pivot_longer(-rowname, names_to = "column", values_to = "value") %>% 
  mutate(rounded_value = round(value, 2))  # Round the correlation values

print(corr_df2)
```

```{r}
g <- ggplot(data = corr_df2, aes(x = rowname, y = column, fill = value)) +
  geom_tile() +
  geom_text(aes(label = rounded_value), fill = "white", size = 3.5) + 
  scale_fill_distiller(palette = "Spectral") +
  labs(title = 'Correlation Matrix', x = '', y = '') +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))

print(g)
```

# Univariate plots:

```{r}
ggplot(df, aes(x = markets, y = popularity)) +
  geom_boxplot()
```

Insight: songs with many genres tend not to be popular

```{r}
ggplot(df, aes(x = view_count, y = like_count, color = genre_level)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black")
```

Insight: idk. Obvious that view count like count correlated

```{r}

```

```{r}

```

```{r}
df$release_date <- as.Date(df$release_date)

ggplot(df, aes(x = release_date, y = popularity, colour = explicitness)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  scale_x_date(
    limits = as.Date(c("2000-01-01", "2024-01-01"))
  ) +
  ylim(0, 100)
```

# Multivariate plots

```{r}
ggplot(df, aes(lexical_richness, after_stat(count), fill = markets)) +
  geom_density(position = "fill")
```

Insight: songs available in very few markets neither tend to have excessive nor limited vocabulary. Perhaps a sign that these songs are rigid af thats why banned???

```{r}
ggplot(df, aes(release_date, after_stat(count), fill = explicitness)) +
  geom_density(position = "fill")
```

Insight: pre 1990 - assume "explicitness" is not an available category. Noticed two "waves" of rise in explicitness. Peaked at 2001. 9-11 attack??? plane crash then stopped having explicit songs

```{r}
ggplot(df, aes(lexical_richness, fill = genre_level)) +
  geom_density(position = "stack")
```

```{r}
density <- ggplot(df, aes(x = sentiment_compound, y = lexical_richness)) +
 geom_density_2d() +
 geom_density_2d_filled(alpha = 0.5) +
 geom_density_2d(linewidth = 0.25, colour = "black")

 density
```

Insight: more happy than sad songs in general. Seems like vocabulary richness of happy songs tend to be slightly less than sad ones. What if we further categorise them?

```{r}
density + facet_wrap(vars(explicitness))
```
```{r}
density + facet_wrap(vars(genre_level))
```

If categorise into explicit or not, non-explicit songs are mostly happy, but explicit ones have a more even spread.

```{r}
ggplot(df, aes(x = popularity, colour = explicitness, fill = explicitness)) +
  geom_density(alpha = 0.2)
```

Insight: explicit songs less popular

```{r}
ggplot(df, aes(x = popularity, colour = markets, fill = markets)) +
  geom_density(alpha = 0.2)
```

Insight: niche market songs tend to be very popular – but has some risks (peaks at low popularity)

```{r}
library(ggradar)
```

```{r}

```

```{r}
ggplot(df, aes(x = duration_seconds, colour = category_number, fill = category_number)) +
  geom_density(alpha = 0.2)
```

```{r}
ggplot(df, aes(x = lexical_richness, y = popularity)) +
 geom_point() +
 geom_density_2d() +
 geom_density_2d_filled(alpha = 0.5) +
 geom_density_2d(linewidth = 0.25, colour = "black")
```

```{r}
ggplot(df, aes(sentiment_compound, fill = markets)) +
  geom_density(position = "stack")
```

```{r}
ggplot(df, aes(x = popularity, y = lexical_richness)) +
  geom_bin2d()
  theme_classic()
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```